---
title: "Week 11: Genome informatics"
author: "Bryn Baxter (PID A69038039)"
format: pdf
---

# Section 1: Porportion on G/G in a population
 
Download CSV file from esemble

We need to read this CSV file 
```{r}
mxl <-read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378 (2).csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl)*100

```

Now lets look at different population. I picked the GBR.

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

Find proportion of G|G

```{r}
table(gbr$Genotype..forward.strand.)/ nrow(gbr)*100
```

This variant that is associated with childhood asthma is more frequent in the GBR population than MKL population. 

Now lets dig into this further.

## Section 4: Population Scale Analysis (HOMEWORK)

How many samples do we have?

```{r}
expr <- read.table("Expression genotype results.txt")
head(expr)
```

```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
library(ggplot2)
```

Lets make a box plot: 
```{r}
ggplot(expr)+ aes(geno,exp, fill=geno)+
  geom_boxplot(notch=T) +
  geom_jitter(width=0.2, alpha=0.2)+
  labs(x="Genotype", y="Expression")
  
```


