---
title: "AlphaFold analysis"
author: "Bryn Baxter PID A69038039"
format: pdf
---

Here we demonstrate how to analyze and make sense of models from AlphaFold. We begin by reading all the model PDB files...

```{r}
library(bio3d)
```

PDB file names of my model 
```{r}
files <- list.files("hiv_monomer_94b5b.result/hiv_monomer_94b5b/", 
                    pattern=".pdb",
                    full.names = T)
files


```

Align and superpose

```{r}
pdbs <- pdbaln(files, fit=T, exefile="msa")

```

RMSD analysis 
```{r}
rd <- rmsd(pdbs, fit=T)
mean(rd)
```

```{r}
source("https://tinyurl.com/newviewngl")
library(NGLVieweR)
#view.pdbs(pdbs)
```

# PCA

```{r}
pc <- pca(pdbs)
plot(pc)
```

#Residue conservation from alignment file 

Alphafold
```{r}
aln_file <- list.files(path="hiv_monomer_94b5b.result/hiv_monomer_94b5b/",
                       pattern=".a3m$",
                        full.names = TRUE)
aln_file
```

```{r}
aln <- read.fasta(aln_file, to.upper = T)
```

```{r}
dim(aln$ali)
```

Score residue conservation:
```{r}
sim <- conserv(aln)
```
```{r}
con <- consensus(aln, cutoff=0.9)
con$seq
```

PLot the conservation along the sequence/structure
```{r}
plot(sim, typ="h")
```
Lets look at these conserved positions in the structure:

```{r}
pdb <- read.pdb(files[1])

#view.pdb(pdb, backgroundColor = "pink", 
#        highlight = atom.select(pdb, resno=25:28), 
#         highlight.style = "spacefill")
```

