---
title: "class 10: Comparative Structure Analysis"
author: "Bryn Baxter (PID A69038039)"
format: html
---

Lets finish off class 10 lab with a "pipeline" for comparative structure analysis. 

```{r}
library(bio3d)

```

```{r}
id <- "1ake_A"


aa <- get.seq(id)
aa

```
Lets search for related sequences in the PDB database 

```{r}
bast <- blast.pdb(aa)
```

Plot an overview of my search results:


```{r}
hits <- plot(bast)
```

Lets see what these "top hits" are:

```{r}
hits$pdb.id
```

We can download all these PDBs to the `path="pdbs"` folder with the `get.pdb()` fucntion:

```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

## Conventional analysis

An RMSD analysis and heatmap:

```{r}
rd <- rmsd(pdbs)
```

```{r}
library(pheatmap)
pheatmap(rd)
```

```{r}
source("https://tinyurl.com/newviewngl")
library(NGLVieweR)
```

```{r}
view.pdbs(pdbs)
```

We can cluster by RMSD and then color our structure display

```{r}
km <- kmeans(rd, centers=3)
mycols <- c("red", "blue", "green")[km$cluster]
```

```{r}
view.pdbs(pdbs, color=mycols)
```

## Principal Component Analysis

```{r}
pc <- pca(pdbs)
plot(pc, col=mycols)
```

```{r}
mktrj(pc, file="pca.pdb")
```

```{r}
p <- read.pdb("pca.pdb", multi=T)
view.pdb(p)
```

```{r}
ref <- read.pdb("1ake")
view.pdb(ref)
```

